---
title: "viz_1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggridges)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USC00519397", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USC00519397 = "Waikiki_HA",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())

```


```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```



```{r}
weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

let make the same plot but different 

```{r}
weather_scatterplot = 
  weather_df %>% 
  drop_na() %>% 
  ggplot(aes(x = tmin, y = tmax)) 

#You can also save the output of ggplot() to an object and modify / print it later. This is often helpful, although it’s not my default approach to making plots.

weather_scatterplot +
  geom_point()
```

Get more fancy 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point()

#OR you can stack, here only aes of color applies to the points 

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,)) + 
  geom_point(aes(color = name)) +
  geom_smooth()

# but here the color applies to all geometires being stacked onto plots & you get smooth lines for all names 

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point() +
  geom_smooth()

# alpha changes color density and se FALSE gets rid of the error bars 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

Make separate panels for each variable through facet wrap 

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax,)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)  + 
  facet_grid(name ~ .) #the "~ ." orients the plots into propper position 
```

comparing new variables, seasonal trend 

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
    facet_grid(. ~ name)
```

#Some quick stuff, C to F conversion 

```{r}
weather_df %>% 
  filter(name == "CentralPark_NY") %>% 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) + 
  geom_smooth(method = "lm", se = FALSE)
```

#change shape of point for better visualization, In these cases geom_hex(), geom_bin2d(), or geom_density2d() can be handy:

```{r}
ggplot(weather_df, aes(x = tmax, y = tmin)) + 
  geom_hex()
```

# Univariate PLots ... histograms, barplots, boxplots, violin 

```{r}
weather_df %>% 
  ggplot(aes(x = tmax)) +
  geom_histogram()

#color is hard to see in historgram so the aes used is "fill"

weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram()
```

More options: Density Plot 

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

More options: Box Plot 

```{r}
weather_df %>% 
  ggplot(aes(y = tmax, x = name, fill = name)) +
  geom_boxplot()
```

More options: Violin Plot, they can be helpful for distribution of many 20+ variables 

```{r}
weather_df %>% 
  ggplot(aes(y = tmax, x = name)) +
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

More options: Ridge Plot, they can be helpful for distribution of many 20+ variables 

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

##saving an embeding plots 

```{r}
weather_scatterplot = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) + 
  geom_smooth( se = FALSE) + 
  facet_grid( . ~ name)

ggsave(
  file = "weather_scatterplot.pdf", 
  plot = weather_scatterplot, 
  width = 8, height = 5
)
```

#you can keep making the figure larger but font will stay same 

```{r fig.width=10}
weather_scatterplot
```

